Investigation of ant cuticle dataset using image texture analysis

### 2022-09-21 paper/main.tex
address minor concerns

The K-means algorithm is one of widely used clustering algorithms in the pattern
recognition \cite{lloyd_least_1982}. It is a single pixel based classification
algorithm for images. Based on the concept of K-means, K-views was developed,
which uses characteristic views for image texture representation
\cite{hung_use_2002}. The K-views algorithm is suitable for classifying image
textures that have basic local patterns repeated in a periodic manner. In
contrast to the K-means, K-views looks at the neighboring pixels for providing
the spatial relationship for classification.  Several variations of the basic
K-views model have been proposed \cite{yang_image_2003, lan_improved_2010}.

For the feasibility study in this ant image datasets, we tested the statistical
clusterings methods and deep learning methods separately. Both K-means and
K-views were used for statistical clustering methods. The experiments may
indicate how well the statistical approach and deep learning networks will do in
this type of textural images. In pre-processing the dataset for the statistical
clustering methods, we apply the Gabor filter to transform the gray-scale ant
image into four Gabor-filtered channels, subtract the gray-scale image from each
channel, and take the absolute value of the resulting images. This
pre-processing will highlight some of characteristic features in textures, such
as edge-like in the image. This step is very useful in feeding the datasets to
the K-views algorithms. The theta $\theta$ parameters in the Gabor filters used
are multiples of $\pi/4$ radians starting from $0$ to $\pi$ radians. The Gabor
filter lambda $\lambda$ parameter is equal to $\pi$.

@@

% \colorbox{green}{(TODO: DRP is best performing model, it should be included in
%     visualization.)}

% \subsection{Clustering segmentation}

% \subsection{t-SNE Visualization}

@@

\newcommand{\subwidth}{0.35\textwidth}
\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
        \caption*{Smooth texture}
        \label{fig:correct_ideal_518}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
        \caption*{Smooth texture}
        \label{fig:correct_ideal_808}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
        \caption*{Rough texture}
        \label{fig:correct_ideal_842}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
        \caption*{Rough texture}
        \label{fig:correct_ideal_1091}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_346}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_1554}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_1694}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
        \caption*{Smooth texture}
        \label{fig:correct_nonideal_388}
    \end{subfigure}
    \caption{GradCAM visualization for correctly classified images from
        fine-tuned ResNet101. The red sections of the heatmap show which features
        had higher activation for the classication.}
    \label{fig:correct_images}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
        \caption*{Rough texture}
        \label{fig:incorrect_ideal_84}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_86}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_177}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_197}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_1}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_22}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_61}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_204}
    \end{subfigure}
    \caption{GradCAM visualization for incorrectly classified images from
        fine-tuned ResNet101. The red sections of the heatmap show which features
        had higher activation for the classication.}
    \label{fig:incorrect_images}
\end{figure}

\newcommand{\segmentedsubwidth}{0.24\textwidth}
\begin{figure}
    \centering
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r103.png}
        \caption*{Rough texture}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
        \caption*{Rough texture}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
        \caption*{Smooth texture}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s107.png}
        \caption*{Smooth texture}
    \end{subfigure}

@@

\FloatBarrier Some classified results from the K-views are shown in Figure
\ref{fig:kviews_segmented}. A training set of 550 patches for background, 455
patches for rough, and 493 patches for smooth are used with patch sizes of 15
$\times$ 15 and 19 $\times$ 19. Hence, the three classes are utilized in the
clustering. These patches are extracted from two rough and two smooth images
with sizes of 608 $\times$ 608. The K-views algorithm converges early although
we set the maximum number of iterations to be 1000. These experimental results
are preliminary due to the minimal number of source images used and future work
will explore a more expansive approach to the problem.

>>>>

The K-means algorithm is one of widely used clustering algorithms in the pattern
recognition \cite{lloyd_least_1982}. It is a single pixel based classification
algorithm for images. Based on the concept of K-means, K-views was developed,
which uses characteristic views for image texture representation
\cite{hung_use_2002}. The K-views algorithm is suitable for classifying image
textures that have basic local patterns repeated in a periodic manner. In
contrast to the K-means, K-views looks at the neighboring pixels for providing
the spatial relationship for classification.

For the feasibility study in this ant image datasets, we tested the statistical
clusterings methods and deep learning methods separately. Both K-means and
K-views were used for statistical clustering methods. The experiments may
indicate how well the statistical approach and deep learning networks will do in
this type of textural images. In pre-processing the dataset for the statistical
clustering methods, we apply the Gabor filter to transform the gray-scale ant
image into four Gabor-filtered channels, subtract the gray-scale image from each
channel, and take the absolute value of the resulting images. This
pre-processing will highlight some of characteristic features in textures, such
as edge-like in the image. This step is very useful in feeding the datasets to
the K-views algorithms. The theta $\theta$ parameters in the Gabor filters used
are multiples of $\pi/4$ radians starting from $0$ to $\pi$ radians. The Gabor
filter lambda $\lambda$ parameter is equal to $\pi$. In order to classify an
image with K-views or K-means we count the number of pixels assigned to each
cluster.

In our experiments, we use $k=3$ clusters; one for background, and the other two
as rough and smooth. To determine if an image is rough or smooth, we choose the
cluster with the majority number of pixels. The number of pixels assigned to the
background cluster are ignored for classification.

@@

@@

\newcommand{\subwidth}{0.35\textwidth}
\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
        \label{fig:correct_ideal_518}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
        \label{fig:correct_ideal_808}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
        \label{fig:correct_ideal_842}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
        \label{fig:correct_ideal_1091}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
        \label{fig:correct_nonideal_346}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
        \label{fig:correct_nonideal_1554}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
        \label{fig:correct_nonideal_1694}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
        \label{fig:correct_nonideal_388}
    \end{subfigure}
    \caption{GradCAM visualization for correctly classified images from
        fine-tuned ResNet101. The red sections of the heatmap show which features
        had higher activation for the classication.}
    \label{fig:correct_images}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
        \label{fig:incorrect_ideal_84}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
        \label{fig:incorrect_ideal_86}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
        \label{fig:incorrect_ideal_177}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
        \label{fig:incorrect_ideal_197}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
        \label{fig:incorrect_nonideal_1}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
        \label{fig:incorrect_nonideal_22}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
        \label{fig:incorrect_nonideal_61}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
        \label{fig:incorrect_nonideal_204}
    \end{subfigure}
    \caption{GradCAM visualization for incorrectly classified images from
        fine-tuned ResNet101. The red sections of the heatmap show which features
        had higher activation for the classication.}
    \label{fig:incorrect_images}
\end{figure}

\newcommand{\segmentedsubwidth}{0.24\textwidth}
\begin{figure}
    \centering
    \begin{subfigure}{\segmentedsubwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \caption*{Rough texture}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \caption*{Smooth texture}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s107.png}
    \end{subfigure}

@@

\FloatBarrier Some classified results from the K-views are shown in Figure
\ref{fig:kviews_segmented}. A training set of 550 patches for background, 455
patches for rough, and 493 patches for smooth are used with patch sizes of 15
$\times$ 15 and 19 $\times$ 19. These patches are extracted from two rough and
two smooth images with sizes of 608 $\times$ 608. The K-views algorithm
converges early although we set the maximum number of iterations to be 1000.
These experimental results are preliminary due to the minimal number of source
images used and future work will explore a more expansive approach to the
problem.
